---
title: "morphine_pk-git"
author: "nbergum"
date: "3/20/2022"
output: github_document
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(car)
library(tidyverse)
library(ggpubr)
library(rstatix)
library(ggthemes)
library(lme4)
```


Data visualization for retina and serum on first day of morphine injections
```{r}
retina_pk <- read.csv("C:\\Users\\nikbe\\Documents\\Morphine Mass Spec\\morphine_pk_retina.csv", fileEncoding = 'UTF-8-BOM')
pk_retina <- retina_pk %>% 
  group_by(ZT,tissue, day, sex, animal) %>% 
  dplyr::summarise(raw_conc = mean(raw_conc), tissue_wt = mean(retina_wt),
                   body_wt= mean(body_wt), dose= mean(dose))
pk_retina$conc_wt <- (pk_retina$raw_conc/pk_retina$tissue_wt)
SumStat_pkr <- dplyr::summarise(group_by(pk_retina, ZT, tissue, day),
n = n(),
mean_conc = mean(raw_conc),
sd_conc = sd(raw_conc),
se_conc = sd_conc/sqrt(n),
mean_conc_wt = mean(conc_wt),
sd_conc_wt = sd(conc_wt),
se_conc_wt = sd_conc_wt/sqrt(n),
log_conc = mean(log(raw_conc)),
sd_conc_log = sd(log(raw_conc)),
se_conc_log = sd_conc_log/sqrt(n),
mean_wt = mean(body_wt),
sd_wt = sd(body_wt),
mean_dose = mean(dose),
sd_dose = sd(dose),
)

serum_pk <- read.csv("C:\\Users\\nikbe\\Documents\\Morphine Mass Spec\\morphine_pk_serum.csv", fileEncoding = 'UTF-8-BOM')
serum_pk
pk_serum <- serum_pk %>% 
  group_by(ZT,tissue, day, sex, animal) %>% 
  dplyr::summarise(raw_conc = mean(raw_conc), tissue_wt = mean(volume),
                   body_wt= mean(body_wt), dose= mean(dose))
pk_serum$conc_wt <- (pk_serum$raw_conc/102.4)
pk_serum
SumStat_pks <- dplyr::summarise(group_by(pk_serum, ZT, tissue, day),
n = n(),
mean_conc = mean(raw_conc),
sd_conc = sd(raw_conc),
se_conc = sd_conc/sqrt(n),
mean_conc_wt = mean(conc_wt),
sd_conc_wt = sd(conc_wt),
se_conc_wt = sd_conc_wt/sqrt(n),
log_conc = mean(log(raw_conc)),
sd_conc_log = sd(log(raw_conc)),
se_conc_log = sd_conc_log/sqrt(n),
mean_wt = mean(body_wt),
sd_wt = sd(body_wt),
mean_dose = mean(dose),
sd_dose = sd(dose),
)

SumStat_pk <- merge(SumStat_pkr, SumStat_pks, all= T)
#View(SumStat_pk) #check files to sample match animals
SumStat_pk1 <- subset(SumStat_pk, day %in% c("1"))
knitr::kable(SumStat_pk1) #only data from day 1 sample collections
write.table(SumStat_pk, file = "C:\\Users\\nikbe\\Documents\\Morphine Mass Spec\\Figures\\SumStats_pk_retina_serum.csv", sep=",", row.names=FALSE)

pk1 <- ggplot(SumStat_pk1, aes(ZT, mean_conc, color=tissue)) + geom_point(stat="identity", position=position_dodge(), size=2) + geom_line(linetype="dotted")  + geom_errorbar(aes(ymin=mean_conc - se_conc, ymax=mean_conc + se_conc, fill = tissue), width=.5) + scale_x_continuous(name="Zeitgeber Time (ZT)", breaks = seq(0,24,2), limits=c(0,24)) + scale_y_continuous(name="Morphine concentration (ng/mL)") + scale_color_manual(values=c("blue", "red4")) + theme_bw() %+replace% theme(text = element_text(size = 14), axis.text=element_text(size=14),  legend.text=element_text(size=14))
pk1

log_pk1 <- ggplot(SumStat_pk1, aes(ZT, log_conc, color=tissue)) + geom_point(stat="identity", position=position_dodge(), size=2) + geom_line(linetype="dotted")  + geom_errorbar(aes(ymin=log_conc - se_conc_log, ymax=log_conc + se_conc_log, fill = tissue), width=.5) + scale_x_continuous(name="Zeitgeber Time (ZT)", breaks = seq(0,24,2), limits=c(0,24)) + scale_y_continuous(name="Log(morphine concentration (ng/mL))") + scale_color_manual(values=c("blue", "red4")) + theme_bw() %+replace% theme(text = element_text(size = 14), axis.text=element_text(size=14),  legend.text=element_text(size=14))
log_pk1

ggsave(file="pk1.svg", plot=pk1, width=6.5, height=4)
ggsave(file="log_pk1.svg", plot=log_pk1, width=6.5, height=4)

pk_retina_serum <- merge(pk_retina, pk_serum, all=T)
pk_retina_serum$log_conc <- log(pk_retina_serum$raw_conc)
pk_retina_serum_day1 <- subset(pk_retina_serum, day %in% c("1"))
retina_serum_1inj<- subset(pk_retina_serum_day1, ZT <12)
pk_retina_serum_day1$ZT <- as.factor(pk_retina_serum_day1$ZT)
pk_retina_serum_day1$tissue <- as.factor(pk_retina_serum_day1$tissue)

#data analysis for day 1 data
lm_day1_logzt <- lmer(log(raw_conc) ~ tissue*ZT + (1|animal), data=pk_retina_serum_day1)
plot(lm_day1_logzt, type=c("p","smooth"), col.line=1)
lattice::qqmath(lm_day1_logzt)
Anova(lm_day1_logzt)
emmeans::emmeans(lm_day1_logzt, pairwise ~ tissue| ZT)
```
                                                                                                                                                                                  Data visualization for retina and serum after single injection
```{r}
SumStat_pk1inj <- dplyr::summarise(group_by(retina_serum_1inj, ZT, tissue, day),
n = n(),
mean_conc = mean(raw_conc),
sd_conc = sd(raw_conc),
se_conc = sd_conc/sqrt(n),
mean_conc_wt = mean(conc_wt),
sd_conc_wt = sd(conc_wt),
se_conc_wt = sd_conc_wt/sqrt(n),
log_conc = mean(log(raw_conc)),
sd_conc_log = sd(log(raw_conc)),
se_conc_log = sd_conc_log/sqrt(n),
log_conc_wt = mean(log(conc_wt)),
sd_conc_wt_log = sd(log(conc_wt)),
)

knitr::kable(SumStat_pk1inj)

serum_retina <- ggplot(SumStat_pk1inj, aes(ZT, mean_conc, color=tissue)) + geom_point(stat="identity", position=position_dodge(), size=2) + geom_line(linetype="dotted")  + geom_errorbar(aes(ymin=mean_conc - se_conc, ymax=mean_conc + se_conc, fill = tissue), width=.5) + scale_x_continuous(name="Zeitgeber Time (ZT)", breaks = seq(0,12,1), limits=c(0,12)) + scale_y_continuous(name="Morphine concentration (ng/mL)", limits = c(0,900)) + scale_color_manual(values=c("blue","red4")) + theme_bw()  %+replace% theme(text = element_text(size = 14), axis.text=element_text(size=14),  legend.text=element_text(size=14))

ggplot(SumStat_pk1inj, aes(ZT, mean_conc_wt, color=tissue)) + geom_point(stat="identity", position=position_dodge(), size=2) + geom_line(linetype="dotted")  + geom_errorbar(aes(ymin=mean_conc_wt - se_conc_wt, ymax=mean_conc_wt + se_conc_wt, fill = tissue), width=.5) + scale_x_continuous(name="Zeitgeber Time (ZT)", breaks = seq(0,12,1), limits=c(0,12)) + scale_y_continuous(name="Morphine concentration per mg tissue (ng/mg)") + scale_color_manual(values=c("blue","red4"))

log_serum_retina <- ggplot(SumStat_pk1inj, aes(ZT, log_conc, color=tissue)) + geom_point(stat="identity", position=position_dodge(), size=2) + geom_line(linetype="dotted")  + geom_errorbar(aes(ymin=log_conc - se_conc_log, ymax=log_conc + se_conc_log, fill = tissue), width=.5) + scale_x_continuous(name="Zeitgeber Time (ZT)", breaks = seq(0,12,1), limits=c(0,12)) + scale_y_continuous(name="log(morphine concentration (ng/mL))", limits = c(0,8)) + scale_color_manual(values=c("blue","red4")) + theme_bw()  %+replace% theme(text = element_text(size = 14), axis.text=element_text(size=14),  legend.text=element_text(size=14))
log_serum_retina 

ggplot(SumStat_pk1inj, aes(ZT, log_conc_wt, color=tissue)) + geom_point(stat="identity", position=position_dodge(), size=2) + geom_line(linetype="dotted")  + geom_errorbar(aes(ymin=log_conc_wt - sd_conc_wt_log, ymax=log_conc_wt + sd_conc_wt_log, fill = tissue), width=.5) + scale_x_continuous(name="Zeitgeber Time (ZT)", breaks = seq(0,12,1), limits=c(0,12)) + scale_y_continuous(name="log(morphine concentration per mg tissue (ng/mg))", limits = c(-4,8)) + scale_color_manual(values=c("blue","red4")) + theme_bw()

ggsave(file="serum_retina.svg", plot=serum_retina, width=6.5, height=4)
ggsave(file="log_serum_retina.svg", plot=log_serum_retina, width=6.5, height=4)
#ggsave("serum_retina.pdf", serum_retina)
#ggsave("log_serum_retina.pdf", log_serum_retina)

ggarrange(serum_retina, log_serum_retina, ncol = 1, nrow = 2)

retina_serum_1inj$ZT <- as.factor(retina_serum_1inj$ZT) 
retina_serum_1inj$log_conc <- log(retina_serum_1inj$raw_conc)
lm_1inj_log <- lm(log(raw_conc) ~ ZT*tissue, data=retina_serum_1inj)
plot(lm_1inj_log)
Anova(lm_1inj_log)
emmeans::emmeans(lm_1inj_log, pairwise ~ tissue| ZT)
shapiro.test(retina_serum_1inj$log_conc)
leveneTest(log(raw_conc) ~ ZT*tissue, data = retina_serum_1inj)

#lmer model accounts for sample matching (the fact that each pair of retina and serum samples were taken from the same animal)
lmer_1inj_log <- lmer(log(raw_conc) ~ ZT*tissue+(1|animal), data=retina_serum_1inj)
plot(lmer_1inj_log, type=c("p","smooth"), col.line=1)
lattice::qqmath(lmer_1inj_log)
Anova(lmer_1inj_log)
emmeans::emmeans(lmer_1inj_log, pairwise ~ tissue| ZT)
```


Data visualization and statistics for retina only after single injection
```{r}
SumStat_pk1injr <- subset(SumStat_pk1inj, tissue %in% c("retina"))
SumStat_pk1injr                   
retina_serum_1inj$ZT <- as.factor(retina_serum_1inj$ZT)

retina <- ggplot(SumStat_pk1injr, aes(ZT, mean_conc, color=tissue)) + geom_point(stat="identity", position=position_dodge(), size=2) + geom_line(linetype="dotted")  + geom_errorbar(aes(ymin=mean_conc - se_conc, ymax=mean_conc + se_conc, fill = tissue), width=.5) + scale_x_continuous(name="Zeitgeber Time (ZT)", breaks = seq(0,12,1), limits=c(0,12)) + scale_y_continuous(name="Morphine concentration (ng/mL)", limits=c(0,900)) + scale_color_manual(values=c("blue")) + theme_bw()  %+replace% theme(text = element_text(size = 14), axis.text=element_text(size=14),  legend.text=element_text(size=14))
retina

log_retina <- ggplot(SumStat_pk1injr, aes(ZT, log_conc, color=tissue)) + geom_point(stat="identity", position=position_dodge(), size=2) + geom_line(linetype="dotted")  + geom_errorbar(aes(ymin=log_conc - se_conc_log, ymax=log_conc + se_conc_log, fill = tissue), width=.5) + scale_x_continuous(name="Zeitgeber Time (ZT)", breaks = seq(0,12,1), limits=c(0,12)) + scale_y_continuous(name="log(morphine concentration (ng/mL))", limits = c(0,8)) + scale_color_manual(values=c("blue")) + theme_bw()  %+replace% theme(text = element_text(size = 14), axis.text=element_text(size=14),  legend.text=element_text(size=14))
log_retina

ggsave(file="retina.svg", plot=retina, width=6.5, height=4)
ggsave(file="log_retina.svg", plot=log_retina, width=6.5, height=4)
#ggsave("serum.pdf", serum)
#ggsave("log_serum.pdf", log_serum)

retina_1injr <- subset(retina_serum_1inj,  tissue %in% c("retina"))
retina_1injr$log_conc <- log(retina_1injr$raw_conc)
lm_1injr <- lm(raw_conc ~ ZT, data=retina_1injr)
plot(lm_1injr)
Anova(lm_1injr)
emoout <- emmeans::emmeans(lm_1injr, pairwise ~ ZT)
emoout
multcomp::cld(emoout$emmeans)
shapiro.test(retina_1injr$raw_conc)
leveneTest(raw_conc ~ ZT, data = retina_1injr)

#log concentrations to try and correct the issues with normality/unequal variance
lm_1injr_log <- lm(log(raw_conc) ~ ZT, data=retina_1injr)
plot(lm_1injr_log)
Anova(lm_1injr_log)
emooutlog<- emmeans::emmeans(lm_1injr_log, pairwise ~ ZT)
emooutlog
multcomp::cld(emooutlog$emmeans)
shapiro.test(retina_1injr$log_conc)
leveneTest(log(raw_conc) ~ ZT, data = retina_1injr)
```

Data visualization and statistics for serum only after single injection
```{r}
SumStat_pk1injs <- subset(SumStat_pk1inj, tissue %in% c("serum"))
SumStat_pk1injs                   
retina_serum_1inj$ZT <- as.factor(retina_serum_1inj$ZT)

serum <- ggplot(SumStat_pk1injs, aes(ZT, mean_conc, color=tissue)) + geom_point(stat="identity", position=position_dodge(), size=2) + geom_line(linetype="dotted")  + geom_errorbar(aes(ymin=mean_conc - se_conc, ymax=mean_conc + se_conc, fill = tissue), width=.5) + scale_x_continuous(name="Zeitgeber Time (ZT)", breaks = seq(0,12,1), limits=c(0,12)) + scale_y_continuous(name="Morphine concentration (ng/mL)", limits = c(0,900)) + scale_color_manual(values="red4")  + theme_bw()  %+replace% theme(text = element_text(size = 14), axis.text=element_text(size=14),  legend.text=element_text(size=14))
serum

log_serum <- ggplot(SumStat_pk1injs, aes(ZT, log_conc, color=tissue)) + geom_point(stat="identity", position=position_dodge(), size=2) + geom_line(linetype="dotted")  + geom_errorbar(aes(ymin=log_conc - se_conc_log, ymax=log_conc + se_conc_log, fill = tissue), width=.5) + scale_x_continuous(name="Zeitgeber Time (ZT)", breaks = seq(0,12,1), limits=c(0,12)) + scale_y_continuous(name="log(morphine concentration (ng/mL))", limits = c(0,8)) + scale_color_manual(values="red4") + theme_bw() %+replace% theme(text = element_text(size = 14), axis.text=element_text(size=14),  legend.text=element_text(size=14))
log_serum

ggsave(file="serum.svg", plot=serum, width=6.5, height=4)
ggsave(file="log_serum.svg", plot=log_serum, width=6.5, height=4)
#ggsave("retina.pdf", retina)
#ggsave("log_retina.pdf", log_retina)

retina_1injs <- subset(retina_serum_1inj,  tissue %in% c("serum"))
retina_1injs$log_conc <- log(retina_1injs$raw_conc)
lm_1injs <- lm(raw_conc ~ ZT, data=retina_1injs)
plot(lm_1injs)

#log concentrations to try and correct the issues with normality/unequal variance
lm_1injs_log <- lm(log(raw_conc) ~ ZT, data=retina_1injs)
plot(lm_1injs_log)
Anova(lm_1injs_log)
emooutlogs <- emmeans::emmeans(lm_1injs_log, pairwise ~ ZT)
multcomp::cld(emooutlogs$emmeans)
shapiro.test(retina_1injs$log_conc)
leveneTest(log(raw_conc) ~ ZT, data = retina_1injs)
```

Comparing morphine concentrations in retina and serum after 1 injection and 2 injections.
```{r}
pk_retina_serum_day1 <- subset(pk_retina_serum, day %in% c("1"))
pk_retina_serum_1_11 <- subset(pk_retina_serum_day1, ZT %in% c("1", "11", "13", "23") )
pk_retina_serum_1_11 <- dplyr::mutate(pk_retina_serum_1_11, injections= ifelse(ZT>12, 2, 1))
pk_retina_serum_1_11 <- dplyr::mutate(pk_retina_serum_1_11, hours = ifelse(ZT == 1 | ZT == 13, 1, 11))
pk_retina_serum_1_11$hours <- as.factor(pk_retina_serum_1_11$hours)
pk_retina_serum_1_11$injections <- as.factor(pk_retina_serum_1_11$injections)
pk_retina_serum_1_11$ZT <- as.factor(pk_retina_serum_1_11$ZT)

#fit model to compare ZTs*tissue against each other for day 1
lm_1_logzt <- lmer(log(raw_conc) ~ tissue*ZT + (1|animal), data=pk_retina_serum_1_11)
plot(lm_1_logzt, type=c("p","smooth"), col.line=1)
lattice::qqmath(lm_1_logzt)
Anova(lm_1_logzt)
emmeans::emmeans(lm_1_logzt, pairwise ~ tissue| ZT)

#fit model to compare injection #, hours post-injections and tissue against each other for day 1
lmer_1_loginj <- lmer(log(raw_conc) ~ tissue*hours*injections + (1|animal), data=pk_retina_serum_1_11)
plot(lmer_1_loginj, type=c("p","smooth"), col.line=1)
lattice::qqmath(lmer_1_loginj)
Anova(lmer_1_loginj)
emmeans::emmeans(lmer_1_loginj, pairwise ~ injections| hours*tissue)

SumStat_pk1_11 <- dplyr::summarise(group_by(pk_retina_serum_1_11, ZT, tissue, day, hours, injections),
n = n(),
mean_conc = mean(raw_conc),
sd_conc = sd(raw_conc),
se_conc = sd_conc/sqrt(n),
log_conc = mean(log(raw_conc)),
sd_conc_log = sd(log(raw_conc)),
se_conc_log = sd_conc_log/sqrt(n),
mean_conc_wt = mean(conc_wt),
sd_conc_wt = sd(conc_wt),
se_conc_wt = sd_conc_wt/sqrt(n),
mean_wt = mean(body_wt),
sd_wt = sd(body_wt),
mean_dose = mean(dose),
sd_dose = sd(dose),
)

SumStat_pk1_11

SumStat_pk1_11$ZT <- as.factor(SumStat_pk1_11$ZT)
SumStat_pk1_11$hours <- as.factor(SumStat_pk1_11$hours)
SumStat_pk1_11$injections <- as.factor(SumStat_pk1_11$injections)
SumStat_pk1_11$tissue <- as.factor(SumStat_pk1_11$tissue)
pk_retina_serum_1_11$ZT <- as.factor(pk_retina_serum_1_11$ZT)

one_11 <- ggplot() + 
  geom_bar(data=SumStat_pk1_11, aes(x=ZT, y=mean_conc, fill=tissue), stat="identity", position=position_dodge(width=0.5), width=0.5) +
  geom_errorbar(data=SumStat_pk1_11, aes(x=ZT, ymin=mean_conc - se_conc, ymax=mean_conc + se_conc, fill=tissue), position=position_dodge(width=0.5), width=.2,) + 
  geom_point(data=pk_retina_serum_1_11, aes(x=ZT, y=raw_conc, fill=tissue), position=position_dodge(width=0.5)) + 
  scale_y_continuous(name="Morphine concentration (ng/mL)") +
  scale_fill_manual(values=c("blue", "red4")) + theme_bw() 

log_one_11 <- ggplot() + 
  geom_bar(data=SumStat_pk1_11, aes(x=ZT, y=log_conc, fill=tissue), stat="identity", position=position_dodge(width=0.5), width=0.5) +
  geom_errorbar(data=SumStat_pk1_11, aes(x=ZT,ymin=log_conc - se_conc_log, ymax=log_conc + se_conc_log, fill=tissue), position=position_dodge(width=0.5), width=.2,) + 
  geom_point(data=pk_retina_serum_1_11, aes(x=ZT, y=log_conc, fill=tissue), position=position_dodge(width=0.5)) + 
  scale_y_continuous(name="log(Morphine concentration (ng/mL))", limits = c(0,10)) +
  scale_fill_manual(values=c("blue", "red4")) +
  xlab("Zeitgeber Time (ZT)") + theme_bw() %+replace% theme(text = element_text(size = 14), axis.text=element_text(size=14),  legend.text=element_text(size=14))

ggplot() + 
  geom_bar(data=SumStat_pk1_11, aes(x=hours, y=mean_conc, fill=injections), stat="identity", position=position_dodge(width=0.5), width=0.5) +
  geom_errorbar(data=SumStat_pk1_11, aes(x=hours,ymin=mean_conc - se_conc, ymax=mean_conc + se_conc, fill=injections), position=position_dodge(width=0.5), width=.2,) + 
  geom_point(data=pk_retina_serum_1_11, aes(x=hours, y=raw_conc, fill=injections), position=position_dodge(width=0.5)) +facet_wrap(~tissue) +
  scale_y_continuous(name="log(Morphine concentration (ng/mL))") + theme_bw() + scale_fill_manual(values=c("grey84", "grey43")) +
  xlab("Hours after each i.p. morphine injection")

one_two <- ggplot() + 
  geom_bar(data=SumStat_pk1_11, aes(x=hours, y=log_conc, fill=injections), stat="identity", position=position_dodge(width=0.5), width=0.5) +
  geom_errorbar(data=SumStat_pk1_11, aes(x=hours,ymin=log_conc - se_conc_log, ymax=log_conc + se_conc_log, fill=injections), position=position_dodge(width=0.5), width=.2,) + 
  geom_point(data=pk_retina_serum_1_11, aes(x=hours, y=log_conc, fill=injections), position=position_dodge(width=0.5)) + facet_wrap(~tissue) +
  scale_y_continuous(name="log(Morphine concentration (ng/mL))", limits = c(0,10)) + theme_bw() + scale_fill_manual(values=c("grey84", "grey43")) +
  xlab("Hours after each i.p. morphine injection")

log_one_two <- ggplot() + 
  geom_bar(data=SumStat_pk1_11, aes(x=hours, y=log_conc, fill=injections), stat="identity", position=position_dodge(width=0.5), width=0.5) +
  geom_errorbar(data=SumStat_pk1_11, aes(x=hours,ymin=log_conc - se_conc_log, ymax=log_conc + se_conc_log, fill=injections), position=position_dodge(width=0.5), width=.2,) + 
  geom_point(data=pk_retina_serum_1_11, aes(x=hours, y=log_conc, fill=injections), position=position_dodge(width=0.5)) + 
  facet_grid(~factor(tissue, levels=c('serum','retina'))) +
  scale_y_continuous(name="log(Morphine concentration (ng/mL))", limits = c(0,10)) + scale_fill_manual(values=c("grey84", "grey43")) +
  xlab("Hours after each i.p. morphine injection") + theme_bw () %+replace% theme(text = element_text(size = 14), axis.text=element_text(size=14),  legend.text=element_text(size=14))

log_one_11
log_one_two

ggsave(file="log_one_11.svg", plot=log_one_11, width=6, height=4)
ggsave(file="log_one_two.svg", plot=log_one_two, width=6, height=4)

#ggsave("log_one_11.pdf", log_one_11)
#ggsave("log_one_two.pdf", log_one_two)
```
Comparing morphine concentrations in retina and serum after 1 , 6 and 13 days of treatment with i.p morphine injections: ZT1 and ZT11.
```{r}

pk_retina_serum_1_11_6 <- subset(pk_retina_serum, ZT %in% c("1", "11"))
pk_retina_serum_1_11_6$ZT <- as.factor(pk_retina_serum_1_11_6$ZT)
pk_retina_serum_1_11_6$day <- as.factor(pk_retina_serum_1_11_6$day)
pk_retina_serum_1_11_6$tissue <- as.factor(pk_retina_serum_1_11_6$tissue)
pk_retina_serum_1_11_6
lm_1_11 <- lmer(log(raw_conc) ~ ZT*tissue*day+(1|animal), data=pk_retina_serum_1_11_6)
plot(lm_1_11, type=c("p","smooth"), col.line=1)
lattice::qqmath(lm_1_11)
Anova(lm_1_11)
emmeans::emmeans(lm_1_11, pairwise ~ tissue | ZT*day)
emmeans::emmeans(lm_1_11, pairwise ~ ZT | day*tissue)
emmeans::emmeans(lm_1_11, pairwise ~ day | ZT*tissue)

SumStat_pk1_11_6 <- dplyr::summarise(group_by(pk_retina_serum_1_11_6, ZT, tissue, day),
n = n(),
mean_conc = mean(raw_conc),
sd_conc = sd(raw_conc),
se_conc = sd_conc/sqrt(n),
log_conc = mean(log(raw_conc)),
sd_conc_log = sd(log(raw_conc)),
se_conc_log = sd_conc_log/sqrt(n),
mean_conc_wt = mean(conc_wt),
sd_conc_wt = sd(conc_wt),
se_conc_wt = sd_conc_wt/sqrt(n),
mean_wt = mean(body_wt),
sd_wt = sd(body_wt),
mean_dose = mean(dose),
sd_dose = sd(dose),
)

SumStat_pk1_11_6$ZT <- as.factor(SumStat_pk1_11_6$ZT)
SumStat_pk1_11_6$day<- as.factor(SumStat_pk1_11_6$day)
SumStat_pk1_11_6$tissue<- as.factor(SumStat_pk1_11_6$tissue)
SumStat_pk1_11_6

ggplot() + 
  geom_bar(data=SumStat_pk1_11_6, aes(x=ZT, y=mean_conc, fill=tissue), stat="identity", position=position_dodge(width=0.5), width=0.5) +
  geom_errorbar(data=SumStat_pk1_11_6, aes(x=ZT,ymin=mean_conc - se_conc, ymax=mean_conc + se_conc, fill=tissue), position=position_dodge(width=0.5), width=.2,) + 
  geom_point(data=pk_retina_serum_1_11_6, aes(x=ZT, y=raw_conc, fill=tissue), position=position_dodge(width=0.5)) + 
  facet_wrap(~day) +
  scale_y_continuous(name="Morphine concentration (ng/mL)") +
  scale_fill_manual(values=c("blue", "red4")) + theme_bw()

ggplot() + 
  geom_bar(data=SumStat_pk1_11_6, aes(x=ZT, y=log_conc, fill=tissue), stat="identity", position=position_dodge(width=0.5), width=0.5) +
  geom_errorbar(data=SumStat_pk1_11_6, aes(x=ZT,ymin=log_conc - se_conc_log, ymax=log_conc + se_conc_log, fill=tissue), position=position_dodge(width=0.5), width=.2,) + 
  geom_point(data=pk_retina_serum_1_11_6, aes(x=ZT, y=log_conc, fill=tissue), position=position_dodge(width=0.5)) + 
  facet_wrap(~day) +
  scale_y_continuous(name="log(Morphine concentration (ng/mL))") +
  scale_fill_manual(values=c("blue", "red4")) + theme_bw()

ggplot() + 
  geom_bar(data=SumStat_pk1_11_6, aes(x=day, y=mean_conc, fill=ZT), stat="identity", position=position_dodge(width=0.5), width=0.5) +
  geom_errorbar(data=SumStat_pk1_11_6, aes(x=day,ymin=mean_conc - se_conc, ymax=mean_conc + se_conc, fill=ZT), position=position_dodge(width=0.5), width=.2,) + 
  geom_point(data=pk_retina_serum_1_11_6, aes(x=day, y=raw_conc, fill=ZT), position=position_dodge(width=0.5)) + 
  facet_grid(~factor(tissue, levels=c('serum','retina'))) +
  scale_y_continuous(name="Morphine concentration (ng/mL)") +
  scale_fill_manual(values=c("slategray1", "slategray4")) + theme_bw()

log_chronic_retina1 <- ggplot() + 
  geom_bar(data=SumStat_pk1_11_6, aes(x=day, y=log_conc, fill=ZT), stat="identity", position=position_dodge(width=0.5), width=0.5) +
  geom_errorbar(data=SumStat_pk1_11_6, aes(x=day,ymin=log_conc - se_conc_log, ymax=log_conc + se_conc_log, fill=ZT), position=position_dodge(width=0.5), width=.2,) + 
  geom_point(data=pk_retina_serum_1_11_6, aes(x=day, y=log_conc, fill=ZT), position=position_dodge(width=0.5)) +  facet_grid(~factor(tissue, levels=c('serum','retina'))) + xlab("Days of morphine treatment") +
  scale_y_continuous(name="log(Morphine concentration (ng/mL))", limits = c(0,10)) +
  scale_fill_manual(values=c("slategray1", "slategray4")) + theme_bw() %+replace% theme(text = element_text(size = 14), axis.text=element_text(size=14),  legend.text=element_text(size=14))

ggplot() + 
  geom_bar(data=SumStat_pk1_11_6, aes(x=ZT, y=mean_conc, fill=day), stat="identity", position=position_dodge(width=0.5), width=0.5) +
  geom_errorbar(data=SumStat_pk1_11_6, aes(x=ZT,ymin=mean_conc - se_conc, ymax=mean_conc + se_conc, fill=day), position=position_dodge(width=0.5), width=.2,) + 
  geom_point(data=pk_retina_serum_1_11_6, aes(x=ZT, y=raw_conc, fill=day), position=position_dodge(width=0.5)) + 
  facet_grid(~factor(tissue, levels=c('serum','retina'))) +
  scale_y_continuous(name="Morphine concentration (ng/mL)") +
  scale_fill_manual(values=c("lightblue","dodgerblue", "blue")) + theme_bw()

log_chronic_retina2 <- ggplot() + 
  geom_bar(data=SumStat_pk1_11_6, aes(x=ZT, y=log_conc, fill=day), stat="identity", position=position_dodge(width=0.5), width=0.5) +
  geom_errorbar(data=SumStat_pk1_11_6, aes(x=ZT,ymin=log_conc - se_conc_log, ymax=log_conc + se_conc_log, fill=day), position=position_dodge(width=0.5), width=.2,) + 
  geom_point(data=pk_retina_serum_1_11_6, aes(x=ZT, y=log_conc, fill=day), position=position_dodge(width=0.5)) + 
  facet_grid(~factor(tissue, levels=c('serum','retina'))) +
  scale_y_continuous(name="log(Morphine concentration (ng/mL))", limits = c(0,10)) +
  scale_fill_manual(values=c("lightblue","dodgerblue", "blue")) + xlab("Zeitgeber Time (ZT)") + theme_bw() %+replace% theme(text = element_text(size = 14), axis.text=element_text(size=14),  legend.text=element_text(size=14))

log_chronic_retina1
log_chronic_retina2

ggsave(file="log_chronic_retina1.svg", plot=log_chronic_retina1, width=6, height=4)
ggsave(file="log_chronic_retina2.svg", plot=log_chronic_retina2, width=6, height=4)

#ggsave("log_chronic_retina1.pdf", log_chronic_retina1)
#ggsave("log_chronic_retina2.pdf", log_chronic_retina2)
```

Comparing morphine concentrations in retina and serum after 1 and 13 days of treatment with i.p morphine injections: ZT2 and ZT9.
```{r}
retina_serum_2_9 <- subset(pk_retina_serum, ZT %in% c("2", "9"))
retina_serum_2_9$day <- as.factor(retina_serum_2_9$day)
retina_serum_2_9$ZT <- as.factor(retina_serum_2_9$ZT)
retina_serum_2_9
lm_2_9 <- lmer(log(raw_conc) ~ ZT*tissue*day+(1|animal), data=retina_serum_2_9)
plot(lm_2_9, type=c("p","smooth"), col.line=1)
lattice::qqmath(lm_2_9)
Anova(lm_2_9)
emmeans::emmeans(lm_2_9, pairwise ~ tissue | ZT*day)
emmeans::emmeans(lm_2_9, pairwise ~ ZT | day*tissue)
emmeans::emmeans(lm_2_9, pairwise ~ day | ZT*tissue)

SumStat_pk2_9 <- subset(SumStat_pk, ZT %in% c("2", "9"))
SumStat_pk2_9$ZT <- as.factor(SumStat_pk2_9$ZT)
SumStat_pk2_9$day<- as.factor(SumStat_pk2_9$day)

ggplot(data=SumStat_pk2_9, aes(x=day, y=mean_conc, fill=ZT)) + geom_bar(stat="identity", position=position_dodge(width=0.5), width=0.5)+ geom_errorbar(aes(ymin=mean_conc - se_conc, ymax=mean_conc + se_conc), position=position_dodge(width=0.5), width=.2) + scale_y_continuous(name="Morphine concentration (ng/mL)") + facet_wrap(~tissue) + scale_fill_manual(values=c("slategray1", "slategray4")) +theme_bw()

ggplot(data=SumStat_pk2_9, aes(x=day, y=log_conc, fill=ZT)) + geom_bar(stat="identity", position=position_dodge(width=0.5), width=0.5)+ geom_errorbar(aes(ymin=log_conc - se_conc_log, ymax=log_conc + se_conc_log), position=position_dodge(width=0.5), width=.2) + scale_y_continuous(name="log(morphine concentration (ng/mL)", limits = c(0,10)) + facet_wrap(~tissue) + theme_bw() + scale_fill_manual(values=c("slategray1", "slategray4")) 

ggplot() + 
  geom_bar(data=SumStat_pk2_9, aes(x=day, y=mean_conc, fill=ZT), stat="identity", position=position_dodge(width=0.5), width=0.5) +
  geom_errorbar(data=SumStat_pk2_9, aes(x=day,ymin=mean_conc - se_conc, ymax=mean_conc + se_conc, fill=ZT), position=position_dodge(width=0.5), width=.2,) +
  geom_point(data=retina_serum_2_9, aes(x=day, y=raw_conc, fill=ZT), position=position_dodge(width=0.5)) +
  facet_grid(~factor(tissue, levels=c('serum','retina'))) +
  scale_y_continuous(name="Morphine concentration (ng/mL)") +
  scale_fill_manual(values=c("slategray1", "slategray4")) + theme_bw()

log_chronic_retina29_1 <-ggplot() + 
  geom_bar(data=SumStat_pk2_9, aes(x=day, y=log_conc, fill=ZT), stat="identity", position=position_dodge(width=0.5), width=0.5) +
  geom_errorbar(data=SumStat_pk2_9, aes(x=day,ymin=log_conc - se_conc_log, ymax=log_conc + se_conc_log, fill=ZT), position=position_dodge(width=0.5), width=.2,) + 
  geom_point(data=retina_serum_2_9, aes(x=day, y=log_conc, fill=ZT), position=position_dodge(width=0.5)) + 
  facet_grid(~factor(tissue, levels=c('serum','retina'))) +
  scale_y_continuous(name="log(Morphine concentration (ng/mL))", limits = c(0,10)) +
  scale_fill_manual(values=c("slategray1", "slategray4")) + theme_bw() + xlab("Day of Morphine treatment") + theme_bw() %+replace% theme(text = element_text(size = 14), axis.text=element_text(size=14),  legend.text=element_text(size=14))

ggplot() + 
  geom_bar(data=SumStat_pk2_9, aes(x=ZT, y=mean_conc, fill=day), stat="identity", position=position_dodge(width=0.5), width=0.5) +
  geom_errorbar(data=SumStat_pk2_9, aes(x=ZT,ymin=mean_conc - se_conc, ymax=mean_conc + se_conc, fill=day), position=position_dodge(width=0.5), width=.2,) +
  geom_point(data=retina_serum_2_9, aes(x=ZT, y=raw_conc, fill=day), position=position_dodge(width=0.5)) +
  facet_grid(~factor(tissue, levels=c('serum','retina'))) +
  scale_y_continuous(name="Morphine concentration (ng/mL)") +
  scale_fill_manual(values=c("lightblue", "blue")) + theme_bw()

log_chronic_retina29_2 <-ggplot() + 
  geom_bar(data=SumStat_pk2_9, aes(x=ZT, y=log_conc, fill=day), stat="identity", position=position_dodge(width=0.5), width=0.5) +
  geom_errorbar(data=SumStat_pk2_9, aes(x=ZT,ymin=log_conc - se_conc_log, ymax=log_conc + se_conc_log, fill=day), position=position_dodge(width=0.5), width=.2,) + 
  geom_point(data=retina_serum_2_9, aes(x=ZT, y=log_conc, fill=day), position=position_dodge(width=0.5)) + 
  facet_grid(~factor(tissue, levels=c('serum','retina'))) +
  scale_y_continuous(name="log(Morphine concentration (ng/mL))", limits = c(0,10)) +
  scale_fill_manual(values=c("lightblue", "blue")) + theme_bw() + xlab("Zeitgeber Time (ZT)") + theme_bw() %+replace% theme(text = element_text(size = 14), axis.text=element_text(size=14),  legend.text=element_text(size=14))

log_chronic_retina29_1
log_chronic_retina29_2

ggsave(file="log_chronic_retina29_1.svg", plot=log_chronic_retina29_1, width=6, height=4)
ggsave(file="log_chronic_retina29_2.svg", plot=log_chronic_retina29_2, width=6, height=4)

#ggsave("log_chronic_retina29_1.pdf", log_chronic_retina29_1)
#ggsave("log_chronic_retina29_2.pdf", log_chronic_retina29_2)
```

Extra data visualizations for morphine concentrations in retina and serum per body weight after 1 , 6 and 13 days of treatment with i.p morphine injections: ZT1 and ZT11.
```{r}
SumStat_pk1_11_1 <- subset(SumStat_pk1_11, ZT %in% c("1", "11"))
SumStat_pk1_11_6$ZT <- as.factor(SumStat_pk1_11_6$ZT)
SumStat_pk1_11_6$day<- as.factor(SumStat_pk1_11_6$day)

ggplot(data=SumStat_pk1_11_6, aes(x=ZT, y=mean_conc_wt, fill=tissue)) + geom_bar(stat="identity", position=position_dodge(width=0.5), width=0.5)+ geom_errorbar(aes(ymin=mean_conc_wt - se_conc_wt, ymax=mean_conc_wt + se_conc_wt), position=position_dodge(width=0.5), width=.2) + scale_y_continuous(name="Morphine concentration per mg tissue (ng/mg)") + facet_wrap(~day) +  scale_fill_manual(values=c("blue", "red4"))

ggplot(data=SumStat_pk1_11_6, aes(x=ZT, y=mean_conc_wt, fill=day)) + geom_bar(stat="identity", position=position_dodge(width=0.5), width=0.5)+ geom_errorbar(aes(ymin=mean_conc_wt - se_conc_wt, ymax=mean_conc_wt + se_conc_wt), position=position_dodge(width=0.5), width=.2) + scale_y_continuous(name="Morphine concentration per mg tissue (ng/mg)") + facet_wrap(~tissue) + scale_fill_manual(values=c("lightblue","dodgerblue", "blue"))

ggplot(data=SumStat_pk1_11_6, aes(x=day, y=mean_conc_wt, fill=ZT)) + geom_bar(stat="identity", position=position_dodge(width=0.5), width=0.5)+ geom_errorbar(aes(ymin=mean_conc_wt - se_conc_wt, ymax=mean_conc_wt + se_conc_wt), position=position_dodge(width=0.5), width=.2) + scale_y_continuous(name="Morphine concentration per mg tissue (ng/mg)") + facet_wrap(~tissue) + scale_fill_manual(values=c("slategray1", "slategray4"))

ggplot(data=SumStat_pk1_11_6, aes(x=day, y=mean_conc_wt, fill=tissue)) + geom_bar(stat="identity", position=position_dodge(width=0.5), width=0.5)+ geom_errorbar(aes(ymin=mean_conc_wt - se_conc_wt, ymax=mean_conc_wt + se_conc_wt), position=position_dodge(width=0.5), width=.2) + scale_y_continuous(name="Morphine concentration per mg tissue (ng/mg)") + facet_wrap(~ZT) +  scale_fill_manual(values=c("blue", "red4"))

```

